Methods

### 2021-10-21 Manuscript/Manuscript_Methods.Rmd
formatting

### Record's acquisition and initial labelling
We built a set of tools to let users automatically search and download citation data from three major scientific databases ("sources"):
Pubmed (<https://pubmed.ncbi.nlm.nih.gov/>), Web Of Science (WOS, <https://apps.webofknowledge.com/>) and the Institute of Electrical and Electronics Engineers (IEEE, <https://ieeexplore.ieee.org/Xplore/home.jsp>).
The framework takes care of authorization management for non-open databases like WOS and IEEE.
It is also possible to download and import records in the framework manually;
this is particularly useful to acquire records from the SCOPUS (<https://www.scopus.com/search/form.uri?display=basic#basic>) and EMBASE databases (<https://www.embase.com/#advancedSearch/default>), for which a comprehensive API interface was not easy to build.
An extra manual search was necessary also for Pubmed, since the API and the web interface have different rule expansion algorithms and return slightly different results [@pubmedUpdate].
A short guide on how to set up the framework for each supported database is available in Supplemental Material S3.\
The acquired records are merged into a single database, resolving duplicates and different formatting between sources.
The records are ordered according to the frequency of the positive query terms (e.g., not preceded by a $NOT$ modifier) in the title and abstract ("simple query ordering").\
The researcher is then asked to label a number of records to create the "initial training set" needed to start the automatic classification.
We suggest manually labelling the first 250 records (see "hyperparameter optimization" later).
The simple query ordering increases the positivity rate in the initial training set [@wallace2010active], which provide higher sensitivity during automatic classification [@chawla2004special].

### Text feature extraction
The collected citation data have a number of fields characterizing a scientific publication.
The framework models the relevance of a record based on the following fields:
title, abstract, authors, keywords, MESH terms [@lipscomb2000medical].
A series of Natural Language Processing (NLP) techniques [@baeza1999modern; @marshall2015systematic; @ananiadou2006text] are employed to transform the textual information in these fields into features for machine learning through a bag-of-words approach [@marshall2015systematic].
The processing of free text fields (title, abstract) includes: tokenization (i.e., extracting the terms), common stopwords (i.e., sentence components bringing no meaning) removal, part-of-speech filtering (only nouns, adjectives, verbs and untagged terms are retained), and lemmatization of the terms (i.e., reduction to their base grammar form).
Text processing for authors, keywords and MESH terms identifies logical units (e.g., author's full names, composite keywords) and extracts them.\
Terms appearing in less than 5% of the labelled documents are removed from negative records.
All terms in the positive set are kept to increase sensitivity at the cost of specificity.\
Some terms tend to co-appear in records (non-consecutive ngrams, nc-ngrams), often carrying a particular meaning when copresent.
To detect nc-ngrams, we generated a word network representation [@rousseau2015graph] posing edges between terms with a cosine similarity in terms of record co-occurrence > 0.5.
We extracted the maximal cliques in the network [@eppstein2010listing] representing highly correlated groups of terms;
These groups of terms are added to the data set as individual features.
Only nc-ngrams comprising a maximum of ten terms are kept.\
A second network is built using a co-occurrence threshold of 0.9.
In this case, the cliques represent terms that always appear together and therefore can be considered redundant (i.e., they do not need to be considered separately).
These terms are merged to increase computation efficiency and reduce overfitting.\
The output is a Document-Term Matrix (DTM), with $N_d$ rows representing the records ($D_i$), $N_t$ terms column for the $t_{field}$ terms (divided by record field) and ${0,1}$ values whether $t_{field} \in D_i$.
We also enriched the DTM with features referencing the number of terms in each field to help the model scale term importance based on the field length.

### Label prediction
We used a Bayesian Additive Regression Trees (BART) machine learning "classification model" [@chipman2010bart] [in the implementation of @kapelner2013bartmachine] to predict the probability of a record of being relevant, given the information coded into the enriched DTM and the initial training set.
We set up the BART model to use 2000 MCMC iterations (after 250 burn-in iterations) and 50 trees;
we used a $k$ value of 2 to regularize extreme prediction and let the model use missing fields in the DTM as features [@kapelner2015prediction].
Positive records are oversampled ten times to increase sensitivity [@batista2004study].\
The output is a posterior predictive distribution (PPD) for each record describing its probability of being relevant (i.e., a positive match).
An ensemble of ten models was fitted to improve prediction stability by averaging the PPD between models [@zhou2021ensemble; @dietterich2000ensemble].\

@@

### Relevant terms' extraction
BART model it is possible to extract the proportion of times a term was used in a posterior tree over the sum of total inclusions of all variables [@kapelner2013bartmachine], as a measure of feature importance for the model.
We extracted the terms, the part of the citation data they were used in, the average "inclusion rate" among the ensemble models (over 10,000 inclusions) and its ratio over the standard deviation of this inclusion (inclusion stability, IS).
For each term we ran a Poisson regression to get its linear association with a positive label and reported it as Relative Risk (RR) with the number of standard errors as significance index (Statistic);
the comparison between the inclusion rate in the BART models and the linear association allows to spot relevant non-linear effects (i.e, the feature is relevant only in association with others).
We reported in the Results only the first (in order of inclusion rate) 15 terms with IS > 1.5 while the first fifty terms regardless of inclusion stability are reported in Supplemental Material S2.

### New search query generation
We created an algorithm that generates a new search query to acquire further relevant publications missed during the first search, possibly at a reasonable cost in specificity (i.e., a higher number of negative results).\
The algorithm encompasses a number of steps:

@@

### Performance evaluation
We trained a simple Bayesian logistic regression (surrogate model) on the reviewed records to evaluate the classification model consistency (see Discussion for the theoretical justification).
The surrogate model uses as predictor the lower bound of the 98% PrI of the records' PPD with weakly regularizing, robust priors for the intercept (Student T with $\nu=3,\mu=0,\sigma=2.5$) and the linear coefficient (Student T with $\nu=3,\mu=0,\sigma=1.5$).\
The quality of the model was evaluated through Bayesian $R^2$ [@gelman2019r], of which we reported the posterior median and 90% Credible Interval [90% CrI].
The $R^2$ also provides an evaluation of the consistency of the original classification model.
Given that this model is conditional only on the BART predictions and not on the DTM, it is characterized by more uncertainty, providing plausible worst-case scenarios.\
The surrogate model is then used to generate the predictive cumulative distribution of the number of missed positive records in the whole dataset.
This distribution allows estimating the expected total posterior "Sensitivity" and the "Work saved over random" (WSoR) of the classification model in the full (unreviewed) dataset.
The WSoR is one minus the number of records to manually label at random to find the same number of positives; this last quantity is estimated through a negative hypergeometric distribution [@chae1993presenting].\
For the number of predicted positive records, the sensitivity and the WSoR, we reported the "truncated 90% PrI" [trunc. 90% PrI], which is the uncertainty interval bounded at the number of observed total positive records (i.e., there cannot be less predicted positive records than observed).

### Hyperparameter evaluation
Our classification algorithm has a number of hyperparameters:

@@

To evaluate the hyperparameter effect of performance, we set up a "grid search" [@claesen2015hyperparameter; @yang2020hyperparameter] on a prelabelled "validation set" made of the first 1200 records from the first session dataset.
See Supplemental Material S1 for the combinations of hyperparameter used and other details on the method.\
The framework tested each hyperparameter combination until four CR iterations with no positive records were returned or the whole dataset got labelled.\
For each combination, a performance score was computed as the product of "Efficiency" (1 minus the ratio of records that required review over the total) and "Sensitivity" (number of positive records found over the total of positives).
We then identified homogeneous "performance clusters" of hyperparameter values and scores using a partition tree.
For the rest of our study, we chose the best cluster of parameter value combinations and, inside the cluster, the best combination according to Sensitivity and Efficiency in this order.

>>>>

### Record's acquisition and initial labelling

We built a set of tools to let users automatically search and download citation data from three major scientific databases ("sources"):
Pubmed (<https://pubmed.ncbi.nlm.nih.gov/>), Web Of Science (WOS, <https://apps.webofknowledge.com/>) and the Institute of Electrical and Electronics Engineers (IEEE, <https://ieeexplore.ieee.org/Xplore/home.jsp>).
The framework takes care of authorization management for non-open databases like WOS and IEEE.
It is also possible to download and import records in the framework manually;
this is particularly useful to acquire records from the SCOPUS (<https://www.scopus.com/search/form.uri?display=basic#basic>) and EMBASE databases (<https://www.embase.com/#advancedSearch/default>), for which a comprehensive API interface was not easy to build.
An extra manual search was necessary also for Pubmed, since the API and the web interface have different rule expansion algorithms and return slightly different results [@pubmedUpdate].
A short guide on how to set up the framework for each supported database is available in Supplemental Material S3.\
The acquired records are merged into a single database, resolving duplicates and different formatting between sources.
The records are ordered according to the frequency of the positive query terms (e.g., not preceded by a $NOT$ modifier) in the title and abstract ("simple query ordering").\
The researcher is then asked to label a number of records to create the "initial training set" needed to start the automatic classification.
We suggest manually labelling the first 250 records (see "hyperparameter optimization" later).
The simple query ordering increases the positivity rate in the initial training set [@wallace2010active], which provide higher sensitivity during automatic classification [@chawla2004special].

### Text feature extraction

The collected citation data have a number of fields characterizing a scientific publication.
The framework models the relevance of a record based on the following fields:
title, abstract, authors, keywords, MESH terms [@lipscomb2000medical].
A series of Natural Language Processing (NLP) techniques [@baeza1999modern; @marshall2015systematic; @ananiadou2006text] are employed to transform the textual information in these fields into features for machine learning through a bag-of-words approach [@marshall2015systematic].
The processing of free text fields (title, abstract) includes: tokenization (i.e., extracting the terms), common stopwords (i.e., sentence components bringing no meaning) removal, part-of-speech filtering (only nouns, adjectives, verbs and untagged terms are retained), and lemmatization of the terms (i.e., reduction to their base grammar form).
Text processing for authors, keywords and MESH terms identifies logical units (e.g., author's full names, composite keywords) and extracts them.\
Terms appearing in less than 5% of the labelled documents are removed from negative records.
All terms in the positive set are kept to increase sensitivity at the cost of specificity.\
Some terms tend to co-appear in records (non-consecutive ngrams, nc-ngrams), often carrying a particular meaning when copresent.
To detect nc-ngrams, we generated a word network representation [@rousseau2015graph] posing edges between terms with a cosine similarity in terms of record co-occurrence > 0.5.
We extracted the maximal cliques in the network [@eppstein2010listing] representing highly correlated groups of terms;
These groups of terms are added to the data set as individual features.
Only nc-ngrams comprising a maximum of ten terms are kept.\
A second network is built using a co-occurrence threshold of 0.9.
In this case, the cliques represent terms that always appear together and therefore can be considered redundant (i.e., they do not need to be considered separately).
These terms are merged to increase computation efficiency and reduce overfitting.\
The output is a Document-Term Matrix (DTM), with $N_d$ rows representing the records ($D_i$), $N_t$ terms column for the $t_{field}$ terms (divided by record field) and ${0,1}$ values whether $t_{field} \in D_i$.
We also enriched the DTM with features referencing the number of terms in each field to help the model scale term importance based on the field length.

### Label prediction

We used a Bayesian Additive Regression Trees (BART) machine learning "classification model" [@chipman2010bart] [in the implementation of @kapelner2013bartmachine] to predict the probability of a record of being relevant, given the information coded into the enriched DTM and the initial training set.
We set up the BART model to use 2000 MCMC iterations (after 250 burn-in iterations) and 50 trees;
we used a $k$ value of 2 to regularize extreme prediction and let the model use missing fields in the DTM as features [@kapelner2015prediction].
Positive records are oversampled ten times to increase sensitivity [@batista2004study].\
The output is a posterior predictive distribution (PPD) for each record describing its probability of being relevant (i.e., a positive match).
An ensemble of ten models was fitted to improve prediction stability by averaging the PPD between models [@zhou2021ensemble; @dietterich2000ensemble].\

@@

### Relevant terms' extraction

BART model it is possible to extract the proportion of times a term was used in a posterior tree over the sum of total inclusions of all variables [@kapelner2013bartmachine], as a measure of feature importance for the model.
We extracted the terms, the part of the citation data they were used in, the average "inclusion rate" among the ensemble models (over 10,000 inclusions) and its ratio over the standard deviation of this inclusion (inclusion stability, IS).
For each term we ran a Poisson regression to get its linear association with a positive label and reported it as Relative Risk (RR) with the number of standard errors as significance index (Statistic);
the comparison between the inclusion rate in the BART models and the linear association allows to spot relevant non-linear effects (i.e, the feature is relevant only in association with others).
We reported in the Results only the first (in order of inclusion rate) 15 terms with IS > 1.5 while the first fifty terms regardless of inclusion stability are reported in Supplemental Material S2.

### New search query generation

We created an algorithm that generates a new search query to acquire further relevant publications missed during the first search, possibly at a reasonable cost in specificity (i.e., a higher number of negative results).\
The algorithm encompasses a number of steps:

@@

### Performance evaluation

We trained a simple Bayesian logistic regression (surrogate model) on the reviewed records to evaluate the classification model consistency (see Discussion for the theoretical justification).
The surrogate model uses as predictor the lower bound of the 98% PrI of the records' PPD with weakly regularizing, robust priors for the intercept (Student T with $\nu=3,\mu=0,\sigma=2.5$) and the linear coefficient (Student T with $\nu=3,\mu=0,\sigma=1.5$).\
The quality of the model was evaluated through Bayesian $R^2$ [@gelman2019r], of which we reported the posterior median and 90% Credible Interval [90% CrI].
The $R^2$ also provides an evaluation of the consistency of the original classification model.
Given that this model is conditional only on the BART predictions and not on the DTM, it is characterized by more uncertainty, providing plausible worst-case scenarios.\
The surrogate model is then used to generate the predictive cumulative distribution of the number of missed positive records in the whole dataset.
This distribution allows estimating the expected total posterior "Sensitivity" and the "Work saved over random" (WSoR) of the classification model in the full (unreviewed) dataset.
The WSoR is one minus the number of records to manually label at random to find the same number of positives; this last quantity is estimated through a negative hypergeometric distribution [@chae1993presenting].\
For the number of predicted positive records, the sensitivity and the WSoR, we reported the "truncated 90% PrI" [trunc. 90% PrI], which is the uncertainty interval bounded at the number of observed total positive records (i.e., there cannot be less predicted positive records than observed).

### Hyperparameter evaluation

Our classification algorithm has a number of hyperparameters:

@@

To evaluate the hyperparameter effect of performance, we set up a "grid search" [@claesen2015hyperparameter; @yang2020hyperparameter] on a prelabelled "validation set" made of the first 1200 records from the first session dataset.
See Supplemental Material S1 for the combinations of hyperparameter used and other details on the method.\
The framework tested each hyperparameter combination until four CR iterations with no positive records were returned or the whole dataset got labelled.\
For each combination, a performance score was computed as the product of "Efficiency" (1 minus the ratio of records that required review over the total) and "Sensitivity" (number of positive records found over the total of positives).
We then identified homogeneous "performance clusters" of scores given hyperparameter values using a partition tree.
For the rest of our study, we chose the best cluster of parameter value combinations and, inside the cluster, the best combination according to Sensitivity and Efficiency, in this order.
